---
title: "ICD-10 PCS codes"
author: "anthonyhung"
date: "2025-07-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Load table

```{r libraries}
library(readr)
library(stringr)
library(tidyverse)
```


```{r data}
# read in full list of 2026 ICD 10 codes from https://www.cms.gov/files/zip/2026-icd-10-pcs-codes-file.zip (7 character code and description)
icd10codes <- read_fwf(
  file = "data/2026-icd-10-pcs-codes-file/icd10pcs_codes_2026.txt",
  fwf_cols(
    code = c(1, 7),
    description = c(9, NA)  # skip 1 space between code and description
  ),
  col_types = cols(.default = "c")
)
```


# Search for the procedures I am interested in and store the codes


## appendectomy = "resection of the appendix"

```{r appendectomy}
appendectomy_codes <- icd10codes %>%
  filter(str_detect(description, regex("resection.*appendix", ignore_case = TRUE))) %>%
  pull(code)

appendectomy_codes_icd9 <- c("47.0","47.01","47.09")

appendectomy_codes_CPT <- c("44950","44955","44960","44970")

appendectomy_codes <- c(appendectomy_codes, appendectomy_codes_icd9, appendectomy_codes_CPT)
```

## cholecystectomy = "resection of the gallbladder"

```{r cholecystectomy}
cholecystectomy_codes <- icd10codes %>%
  filter(str_detect(description, regex("resection.*gallbladder", ignore_case = TRUE))) %>%
  pull(code)

cholecystectomy_codes_icd9 <- c("51.2","51.22","51.23")

cholecystectomy_codes_CPT <- c("47562","47563","47564","47600","47605","47610","47612","47620")

cholecystectomy_codes <- c(cholecystectomy_codes, cholecystectomy_codes_icd9, cholecystectomy_codes_CPT)
```

## colon resection (there are many parts)

```{r colon resection}
large_intestine_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Large intestine", ignore_case = TRUE))) %>%
  pull(code)

cecum_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Cecum", ignore_case = TRUE))) %>%
  pull(code)

ascending_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Ascending Colon", ignore_case = TRUE))) %>%
  pull(code)

transverse_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Transverse Colon", ignore_case = TRUE))) %>%
  pull(code)

descending_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Descending Colon", ignore_case = TRUE))) %>%
  pull(code)

sigmoid_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Sigmoid Colon", ignore_case = TRUE))) %>%
  pull(code)

left_colon_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Left Large Intestine", ignore_case = TRUE))) %>%
  pull(code)

right_colon_codes <- icd10codes %>%
  filter(str_detect(description, regex("^Resection of Right Large Intestine", ignore_case = TRUE))) %>%
  pull(code)

colon_codes <- c(large_intestine_codes,
                 cecum_codes,
                 ascending_codes,
                 transverse_codes,
                 descending_codes,
                 sigmoid_codes,
                 left_colon_codes,
                 right_colon_codes)

colon_codes_icd9 <- c("17.3","17.31","17.32","17.33","17.34","17.35","17.36","17.39","45.7","45.71","45.72","45.73","45.74","45.75","45.76","45.79","45.8","45.81","45.82","45.83")

colon_codes_CPT <- c("44144","44157","44158","44160","44204","44205","44207","44208","45402")

colon_codes <- c(colon_codes, colon_codes_icd9, colon_codes_CPT)
```

## Gastric bypass (stomach to Jejunum or ileum)

```{r gastric bypass}
gastric_bypass_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Bypass Stomach to (Jejunum|Ileum)", ignore_case = TRUE))
  ) %>%
  pull(code)

gastric_bypass_codes_icd9 <- c("43.7","43.81","44.31","44.32","44.38","44.39")

gastric_bypass_codes_CPT <- c("43644","43645","43845","43846","43847")

gastric_bypass_codes <- c(gastric_bypass_codes, gastric_bypass_codes_icd9, gastric_bypass_codes_CPT)
```

## Anterior or posterior spinal arthrodesis

```{r spinal fusions}
spinal_fusion_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Fusion of .*vertebral joint", ignore_case = TRUE)) &
    str_detect(description, regex("anterior|posterior", ignore_case = TRUE))
  ) %>%
  pull(code)

spinal_fusion_codes_icd9 <- c("81.00","81.01","81.02","81.03","81.04","81.05","81.06","81.07","81.08","81.30","81.31","81.32","81.33","81.34","81.35","81.36","81.37","81.38","81.39","81.62","81.63","81.64","81.65","81.66")

spinal_fusion_codes_CPT <- c("22548","22551","22552","22554","22556","22558","22585","22586","22590","22595","22600","22610","22612","22614","22630","22632","22633","22634","22800","22802","22804","22808","22810","22812")

spinal_fusion_codes <- c(spinal_fusion_codes, spinal_fusion_codes_icd9, spinal_fusion_codes_CPT)
```

## Total knee replacement

```{r TKA}
knee_replacement_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Replacement of (Right|Left) Knee Joint", ignore_case = TRUE))
  ) %>%
  pull(code)

knee_replacement_codes_icd9 <- "81.54"

knee_replacement_codes_CPT <- c("27447","27487")

knee_replacement_codes <- c(knee_replacement_codes, knee_replacement_codes_icd9, knee_replacement_codes_CPT)
```

## Total hip replacement

```{r THA}
hip_replacement_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Replacement of (Right|Left) Hip Joint", ignore_case = TRUE))
  ) %>%
  pull(code)

hip_replacement_codes_icd9 <- "81.51"

hip_replacement_codes_CPT <- c("27130","27132","27134")

hip_replacement_codes <- c(hip_replacement_codes, hip_replacement_codes_icd9, hip_replacement_codes_CPT)
```

## Craniotomy for brain tumor (open approach)

```{r craniotomy}
craniotomy_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Excision of Brain, Open Approach", ignore_case = TRUE))
  ) %>%
  pull(code)

craniotomy_codes_icd9 <- c("01.24","01.31","01.32","01.39","01.53")

craniotomy_codes_CPT <- c("61500","61510","61512","61518","61519","61520","61521","61522","61524","61526","61530","61545","61546")

craniotomy_codes <- c(craniotomy_codes, craniotomy_codes_icd9, craniotomy_codes_CPT)
```

## Total thyroidectomy

```{r thyroidectomy}
thyroidectomy_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Resection of Thyroid Gland,", ignore_case = TRUE))
  ) %>%
  pull(code)

thyroidectomy_codes_icd9 <- "06.4"

thyroidectomy_codes_CPT <- c("60240","60252","60254","60260","60270","60271")

thyroidectomy_codes <- c(thyroidectomy_codes, thyroidectomy_codes_icd9, thyroidectomy_codes_CPT)
```

## Femoral-popliteal bypass

```{r fem-pop}
fem_pop_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Bypass (Right|Left) Femoral Artery.* to Popliteal Artery", ignore_case = TRUE))
  ) %>%
  pull(code)

fem_pop_codes_icd9 <- "39.29"

fem_pop_codes_CPT <- c("35556","35583","35656")

fem_pop_codes <- c(fem_pop_codes, fem_pop_codes_icd9, fem_pop_codes_CPT)
```

## Lung resection (note: this omits trachea and carina)

```{r lung resection}
lung_resection_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Resection of.* Lung|^Resection of.* Bronchus", ignore_case = TRUE))
  ) %>%
  pull(code)

lung_resection_codes_icd9 <- c("32.30","32.39","32.41","32.49","32.50","32.59")

lung_resection_codes_CPT <- c("32491","32507","32668")

lung_resection_codes <- c(lung_resection_codes, lung_resection_codes_icd9, lung_resection_codes_CPT)
```

## Open repair femoral neck/shaft fracture (Not replacement, "upper or lower femur", "open")

```{r femur_repair resection}
femur_repair_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^Repair (Right|Left) (Upper|lower) Femur, Open", ignore_case = TRUE))
  ) %>%
  pull(code)

femur_repair_codes_icd9 <- c("79.25","79.35","79.55")

femur_repair_codes_CPT <- c("27170","27176","27177","27178","27179","27181","27248","27470","27472")

femur_repair_codes <- c(femur_repair_codes, femur_repair_codes_icd9, femur_repair_codes_CPT)
```

## AAA repair ("supplement, repair, or replace")

```{r AAA}
AAA_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("^(Supplement|Repair|Replacement|Restriction) of Abdominal Aorta", ignore_case = TRUE))
  ) %>%
  pull(code)

AAA_codes_icd9 <- "39.52"

AAA_codes_CPT <- c("34701","34702","34703","34704","34705","34706","34830","34831","34832","34841","34842","34843","34844","34845","34846","34847","34848","35081","35082","35091","35092","35102","35103")

AAA_codes <- c(AAA_codes, AAA_codes_icd9, AAA_codes_CPT)
```

## Neck dissection (excision or resection of neck lymphatics)

```{r neck}
neck_codes <- icd10codes %>%
  filter(
    str_detect(description, regex("(Resection|Excision).* neck lymphatic", ignore_case = TRUE))
  ) %>%
  pull(code)

neck_codes_icd9 <- c("40.40","40.41","40.42")

neck_codes_CPT <- c("31365","31368","38542","38720","38724","41135","41145","41155","42426","69155")

neck_codes <- c(neck_codes, neck_codes_icd9, neck_codes_CPT)
```

## Make dataframe of codes/descriptions
```{r export}
code_list <- list(
  AAA = AAA_codes,
  appendectomy = appendectomy_codes,
  cholecystectomy = cholecystectomy_codes,
  colon = colon_codes,
  craniotomy = craniotomy_codes,
  fem_pop = fem_pop_codes,
  femur = femur_repair_codes,
  gastric_bypass = gastric_bypass_codes,
  hip = hip_replacement_codes,
  knee = knee_replacement_codes,
  lung = lung_resection_codes,
  neck = neck_codes,
  spine = spinal_fusion_codes,
  thyroidectomy = thyroidectomy_codes
)

# Convert to tibble
df <- enframe(code_list, name = "group", value = "codes") %>%
  unnest(codes)

df$codes <- gsub("\\.", "", df$codes)

write.csv(df, "output/surgery_codes.csv", row.names = FALSE)
```


