---
title: "noncardiacSGY_code_lookup"
author: "anthonyhung"
date: "2025-10-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(dplyr)
library(stringr)
library(readr)
```

# ICD-10-PCS

## ICD-10-PCS codes from the HCUP CCSR categories
- included amongst the categories are 
appendectomy
cholecystectomy
gastric bypass - "Gastro-jejunal bypass (including bariatric)"
colon resection - "colectomy"
anterior or posterior spinal arthrodesis - "Spine fusion"
craniotomy for brain tumor - ?"CNS excision procedures"
total knee replacement - ?"Knee arthroplasty"
total hip replacement - ?"Hip arthroplasty"
open repair of femoral neck or shaft fracture - ?"Femur fixation"
total thyroidectomy 
lung resection - "Lung, pleura, or diaphragm resection (open and thoracoscopic)"

```{r}
ICD10_CSS <- read_csv("data/PRCCSR_v2025-1/PRCCSR-ICD10_mapping.csv")

ICD10_CSS_appendectomy <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Appendectomy") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_chole <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Cholecystectomy") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_gastricBypass <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Gastro-jejunal bypass (including bariatric)") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_colectomy <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Colectomy") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_spineFusion <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Spine fusion") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_craniotomy <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "CNS excision procedures") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_TKA <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Knee arthroplasty") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_THA <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Hip arthroplasty") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_femur <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Femur fixation") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_thyroid <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Thyroidectomy") %>% 
  pull(`ICD-10-PCS Code`)

ICD10_CSS_lung <- ICD10_CSS %>% 
  filter(`CCSR Category Description` == "Lung, pleura, or diaphragm resection (open and thoracoscopic)") %>% 
  pull(`ICD-10-PCS Code`)

#Stack all the vectors
ICD10_CSS_List <- stack(list(
  ICD10_CSS_appendectomy = ICD10_CSS_appendectomy,
  ICD10_CSS_chole = ICD10_CSS_chole,
  ICD10_CSS_gastricBypass = ICD10_CSS_gastricBypass,
  ICD10_CSS_colectomy = ICD10_CSS_colectomy,
  ICD10_CSS_spineFusion = ICD10_CSS_spineFusion,
  ICD10_CSS_craniotomy = ICD10_CSS_craniotomy,
  ICD10_CSS_TKA = ICD10_CSS_TKA,
  ICD10_CSS_THA = ICD10_CSS_THA,
  ICD10_CSS_femur = ICD10_CSS_femur,
  ICD10_CSS_thyroid = ICD10_CSS_thyroid,
  ICD10_CSS_lung = ICD10_CSS_lung
))
names(ICD10_CSS_List) <- c("code", "Type")
```

line-by-line file

```{r}
# read in full list of 2026 ICD 10 codes from https://www.cms.gov/files/zip/2026-icd-10-pcs-codes-file.zip (7 character code and description)
icd10codes <- read_fwf(
  file = "data/2026-icd-10-pcs-codes-file/icd10pcs_codes_2026.txt",
  fwf_cols(
    code = c(1, 7),
    description = c(9, NA)  # skip 1 space between code and description
  ),
  col_types = cols(.default = "c")
)
head(icd10codes)

# make into tibble for joining
ICD10_CSS_List <- tibble(ICD10_CSS_List)

# join with icd10codes on the code
noncardiacSGYlookup <- ICD10_CSS_List %>%
  left_join(icd10codes, by = c("code")) %>%
  rename(`Number / Range` = code,
         `Full Description` = description) %>%
  mutate(
    Approach = str_extract(`Full Description`,
                           "(Open Approach|Percutaneous Endoscopic Approach|Percutaneous Approach|Natural or Artificial Opening Endoscopic|External Approach|Natural or Artificial Opening With Percutaneous Endoscopic Assistance)")
  )
```


###todo

```{r}
# Function to place the ICD-10-PCS codes into the correct body part/approach category
classify_codes <- function(codes) {
  
  df <- data.frame(
    code = codes,
    body_part = substr(codes, 4, 4),
    approach = substr(codes, 5, 5),
    stringsAsFactors = FALSE
  )
  
  # Valves: map by body part character position
  valve_map <- c("F" = "Aortic", "G" = "Mitral", "H" = "Pulmonic", "J" = "Tricuspid")
  df$valve <- valve_map[df$body_part]
  
  df$type <- case_when(
    df$approach == "0" ~ "OPEN",
    df$approach %in% c("3", "4") ~ "PERC",
    TRUE ~ NA
  )
  
  df$category <- ifelse(is.na(df$category) & !is.na(df$valve),
                        paste0(df$valve, df$type),
                        df$category)
  
  # Collapse to summary dataframe
  summary_df <- df %>%
    filter(!is.na(category)) %>%
    group_by(category) %>%
    summarise(codes = paste(unique(code), collapse = ", "),
              .groups = "drop")
  
  return(summary_df)
}
```


# CPT codes

## CPT codes from the HCUP CCSR categories
- included amongst the categories are 

appendectomy
cholecystectomy - "Cholecystectomy and common duct exploration"
gastric bypass - "Gastric bypass and volume reduction"
anterior or posterior spinal arthrodesis - "spinal fusion"
total knee replacement - "Arthroplasty knee"
total hip replacement - "Hip replacement, total and partial"
open repair of femoral neck or shaft fracture - "Treatment, fracture or dislocation of hip and femur"
total thyroidectomy - "Thyroidectomy, partial or complete"
lung resection - "Lobectomy or pneumonectomy"


```{r}
CPT_CSS <- read_csv("data/CCS-Services_Procedures_v2025-1/CCS_Services_Procedures_v2025-1_052425.csv")

CPT_CSS_appendectomy <- CPT_CSS %>% 
  filter(`CCS Label`  == "Appendectomy") %>% 
  pull(`Code Range`)

CPT_CSS_chole <- CPT_CSS %>% 
  filter(`CCS Label`  == "Cholecystectomy and common duct exploration") %>% 
  pull(`Code Range`)

CPT_CSS_gastricBypass <- CPT_CSS %>% 
  filter(`CCS Label`  == "Gastric bypass and volume reduction") %>% 
  pull(`Code Range`)

CPT_CSS_spineFusion <- CPT_CSS %>% 
  filter(`CCS Label`  == "Spinal fusion") %>% 
  pull(`Code Range`)

CPT_CSS_TKA <- CPT_CSS %>% 
  filter(`CCS Label`  == "Arthroplasty knee") %>% 
  pull(`Code Range`)

CPT_CSS_THA <- CPT_CSS %>% 
  filter(`CCS Label`  == "Hip replacement, total and partial") %>% 
  pull(`Code Range`)

CPT_CSS_femur <- CPT_CSS %>% 
  filter(`CCS Label`  == "Treatment, fracture or dislocation of hip and femur") %>% 
  pull(`Code Range`)

CPT_CSS_thyroid <- CPT_CSS %>% 
  filter(`CCS Label`  == "Thyroidectomy, partial or complete") %>% 
  pull(`Code Range`)

CPT_CSS_lung <- CPT_CSS %>% 
  filter(`CCS Label`  == "Lobectomy or pneumonectomy") %>% 
  pull(`Code Range`)


#Stack all the vectors
CPT_CSS_List <- stack(list(
  CPT_CSS_appendectomy = CPT_CSS_appendectomy,
  CPT_CSS_chole = CPT_CSS_chole,
  CPT_CSS_gastricBypass = CPT_CSS_gastricBypass,
  CPT_CSS_spineFusion = CPT_CSS_spineFusion,
  CPT_CSS_TKA = CPT_CSS_TKA,
  CPT_CSS_THA = CPT_CSS_THA,
  CPT_CSS_femur = CPT_CSS_femur,
  CPT_CSS_thyroid = CPT_CSS_thyroid,
  CPT_CSS_lung = CPT_CSS_lung
))
names(CPT_CSS_List) <- c("code", "Type")

CPT_CSS_List$code <- gsub("^'?(\\w+)-\\1'?$", "'\\1'", CPT_CSS_List$code)
```




