---
title: "surgCodes"
author: "anthonyhung"
date: "2025-07-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r}
library(dplyr)
```


## Identify ICD codes corresponding to surgeries and categories of surgeries



```{r}
PClass <- read.csv("data/PClassR_v2023-1/PClassR_v2023-1.csv",
                   skip = 1,
                   header = TRUE)
PClass_surg <- PClass %>%
  filter(X.PROCEDURE.CLASS. == 3 | X.PROCEDURE.CLASS. == 4) # identify "Major" class diagnostic or therapeutic procedures (likely to be OR operations)

#Put codes into categories based on PRCCSR
PRCCSR_categories_file <- read.csv("data/PRCCSR_v2023-1/PRCCSR_categories.csv") #I made this file from the reference excel file's panel
PRCCSR <- read.csv("data/PRCCSR_v2023-1/PRCCSR_v2023-1.CSV")

#join datasets
joined_codes <- merge(PClass_surg, PRCCSR, by.x = "X.ICD.10.PCS.CODE.", by.y = "X.ICD.10.PCS.")

dim(joined_codes) #56454 codes qualify
table(joined_codes$X.CLINICAL.DOMAIN.)

write.csv(joined_codes$X.ICD.10.PCS.CODE., "~/Desktop/PCScodes.csv")
```


Surgeries from etherington et al

Femoral-popliteal bypass
AAA repair
Appendectomy
Cholecystectomy
Gastric bypass
Colon resection
Anterior or posterior spinal arthrodesis
Craniotomy for a brain tumor
Total knee replacement
Total hip replacement
Open repair of femoral neck or shaft fracture
Total thyroidectomy
Neck dissection
Lung resection


